MA_Results

### 2023-04-19 Markdown/Manuscript/Child_RMDs/MS_Imp_Guidelines.Rmd
first pass and updated bib

Based on published guidance on computational modelling in health economics and other disciplines as well as our own experience, we have identified 20 standards that we believe are important for implementing a MOSHEM that is accountable (seven standards), reusable (nine standards) and updatable (four standards). The standards are specified in Table \@ref(tab:timelygls) and are discussed below.

@@

### Standards for an accountable MOSHEM
Guidance on transparency in health economic modelling published over ten years ago [@Eddy2012] made recommendations on documenting models but notably did not include recommendations on sharing model code and data. However, more recent and multidisciplinary healthcare modelling guidance [@Erdemir2020] recommends using existing digital repository services to these types of digital model artefacts. Some types of repository such as GitHub [@github2007] provide tools for disseminating work in progress code and providing highly transparent records of the complete development history and individual authorship contributions of a software project, while others such as Zenodo [@Zenodo2013] and Dataverse [@Dataverse2007] provide persistent storage solutions that generate a Digital Object Identifier (DOI) for each code and data collection.

Model code and data should be clearly documented, potentially with different versions for technical and non-technical users [@Eddy2012]. Consistent use of meaningful naming conventions when authoring code is recommended [@Wilson_2017; @Alarid2019]. Code can be made easier to follow by using the practices of abstraction [@8717448], where only simple, high level commands are routinely exposed to reviewers, and polymorphism [@7181447], where the same command (e.g. "simulate") can be reused to implement different algorithms of the same type. Programs to implement model analyses can be made comprehensible to even non-technical users through the use of literate programming techniques and tools like RMarkdown [@xie2018r] that integrate computer code with plain English descriptions.

An essential component of quality assuring health economic models is verification - ensuring that calculations are correct and consistent with model specifications [@techver2019]. One useful concept for informing model users about the extensiveness of verification checks is code coverage [@ERICWONG2010188] - the proportion of model code that has been explicitly tested. <!-- In R, the testthat [@testthat2011] and covr [@covr2020] tools can be used in conjunction with GitHub to define tests and report coverage metrics. -->Transcription errors - mistakes introduced when transferring data between sources, models and reports - are very common in health economic models [@Radeva2020]. The risk of these errors might be lower if there was full transparency across all steps in a study workflow. Scientific computing tools now make it relatively straightforward to author programs that reproducibly execute all steps in data ingest, processing and reporting [@Wilson_2017].

Code and data should be distributed with tools that make it easy for potential users to appropriately cite each model artefact.

### Standards for a reusable MOSHEM
```{r echo = F}
gpl_pc_dbl <- get_license_share("GPL")
```

To make model code and data widely re-usable by others, it is important to provide users with appropriate and explicit permissions. In the context of open source models, there are two broad categories of licensing options. Some guidance strongly recommends the use of permissive licensing [@Wilson_2017] that provides users with great flexibility as to the purposes (including commercial) for which the content could be re-used. An alternative approach is to use copyleft licenses [@copyleft2022] that can require content users to distribute any derivative works they create under similar open source arrangements. For code, it may be appropriate to adopt the prevailing open source licensing practice within the programming language being used. For data, it may not be sufficient to simply choose between a permissive license like the Public Domain Dedication (CC0) [@cc02022] or a copyleft option such as the Attribution-Share Alike (CC-BY-SA) [@bysa2022]. In addition to ensuring that data is ethically appropriate for disseminate in open access repositories, responsible custodianship of some de-identified or aggregated data may involve using or adapting template terms of use [@sampleterms2022] which have a number of ethical clauses (for example, prohibiting efforts to re-identify research participants).

Storing model code and data in separate files and locations (as opposed to hard coding - embedding data into source code) can make it easier to apply models to different decision contexts and to selectively restrict access to data that are confidential, while disseminating all other model artefacts. Clear distinctions should be made between model modules (code that defines abstract data structures and the algorithms that can be applied to data described by these structures), model datasets (digital information such as parameter values, unit records, etc) and model analyses (code that links model datasets to model modules and specifies the algorithms to apply to data associated with each module).

The software development practice of encapsulation [@8717448] can be used to help ensure that model modules continue to work as intended when they are combined [@ready4oop2022]. In some cases, combining modules may mean new versions of modules have to be created to better account for interaction effects. The concept of inheritance [@8717448] can be used to write code that efficiently achieves this objective as well as to facilitate selective editing of modules when transferring models to different decision contexts [@ready4oop2022]. Writing algorithms as collections of functions (short, self-contained and reusable software routines that each perform a discrete task) is recommended as good practice for scientific computing [@Wilson_2017]. Functions to implement model algorithms can be associated with data structures (also known as a class) via a special type of function called a method. Model modules of a similar type or purpose can be efficiently distributed and documented by bundling them as code libraries. It is good practice to make available test or toy data to demonstrate the use of model algorithms [@Wilson_2017].

Statistical models are a common output of health economic evaluations, but they are often not reported in a format that enables others to confidently and reliably re-use them for out of sample prediction [@Kearns2013]. Open source approaches can help address this by disseminating code artefacts that enable easy and appropriate use of a statistical model to make predictions with new data. However, great care must be exercised when doing so if models are derived from data on human subjects as some software artefacts by default contain a copy of the source dataset. Such dataset copies must therefore be replaced (for example, with synthetic data) and the amended artefact's predictive performance then retested before any public release. Another way to make MOSHEMs easier to use is to develop simple user-interfaces for non-technical users.

### Standards for an updatable MOSHEM
To avoid MOSHEMs going stale - losing validity and usefulness with time - they should be routinely updated.  Each update of code and data should be uniquely identifiable and retrievable, a goal that can be facilitated by use of version control tools [@Erdemir2020] such as [@git20XX].

Potential users of model code and data should be able to easily identify which version is most appropriate to their needs. Using semantic versioning [@semver20xx] conventions can signal the potential importance of an update to users of model code and data. Also informative is to clearly label the type of code hosted in a repository as either:

- "development" (typically the most comprehensive and up to date source code, but potentially not yet sufficiently well documented and tested for potential users to apply to purposes other than testing);

- "production" (code that has been released with a view to it being applied to its stated purposes and therefore typically required to meet defined documentation and testing standards); or

- "archive" (permanent copies of code at key milestones in its development - potentially useful if seeking to use older versions of code to reproduce analyses).

Continuous integration [@CI2017] tools can help verify that each code update passes multiple quality tests. Finally, using deprecation conventions that take an informative and staged approach to retiring old code and data reduces the risk that model revisions have unintended consequences on third party users.

>>>>

We considered prior literature on modelling practice, our own professional experience and the needs of our project, to identify three core ethical responsibilities of CHEM developers, three attributes of CHEMs that enable fulfilment of these responsibilities and six criteria (two for each model attribute) for assessing ethical CHEM implementations.

## Ethical responsibilities of CHEM developers 
We believe that health economists have ethical responsibilities relating to the social acceptability, adequacy for purpose and beneficial impact of their computational models.

Misalignment between the values of model developers and those of the population groups affected by decisions based on their models presents significant ethical risks [@thompson2022escape; @duckett2022journey]. The value judgments of model developers influence the assumptions, selection of model features and standards for evidence that shape the CHEM development process [@HARVARD2020112975]. These value judgments are rarely made explicit, omissions that may lead to socially unacceptable policy recommendations [@duckett2022journey]. For example, to reduce the risk of inequitable policy implementation, it may be important for of a model to predict the benefits of harms of an intervention for different sub-populations [@10.3389/fpubh.2017.00068], but model developers may prefer to allocate a project budget to implementing other model features [@thompson2022escape].

Health economists have duties both to take sufficient care that a CHEM is adequate for the explicit purpose for which it was developed and to provide potential third party users with the means of assessing its adequacy for their proposed purposes [@Erdemir2020; @Feenstra2022; @thompson2019escape; @10.3389/fpubh.2017.00068]. However, it is common for CHEMs to have serious methodological flaws [@carletto_zanuzzi_sammarco_russo_2020; @WONDER2015467]; insufficient validation [@Ghabri2019; @kolovos2017model; @haji2013model], poor reproducibility [@Jalali2021; @McManus2019; @Bermejo2017]; and undeclared errors [@Radeva2020].

Even an acceptable and adequate CHEM will have limited beneficial impact if it not much used, if it is mis-used or when its acceptability and adequacy rapidly decay. The scientific goals of generalisability (application without adaptation) and transferability (selective reuse and/or modification of model components) [@RN39] are advanced when models are appropriately reused in new contexts. Reuse of models as components of other models can also make model implementation more efficient [@Arnold2010; @garcia2021cost].

Common barriers to model re-use include commercial and legal considerations [@Feenstra2022; @Pouwels2022], as well as challenges related to model transferability across jurisdictions [@garcia2021cost]. The temporal window for valid application of CHEMs is often limited by implementation choices that rarely facilitate routine updates [@Sampson2017]. Without ongoing maintenance, a model may become less reliable with time [@garcia2021cost], deterioration that model users may be unaware of, and has a growing risk of being deployed for purposes for which it is poorly suited [@calder2018computational].

##	Criteria for assessing ethical CHEM implementation
The acceptability, adequacy and public benefit responsibilities for CHEM developers are easier to state than to measure.  It may therefore be pragmatic for assessment criteria for ethical modelling practice to instead be based on measurable attributes of the models themselves. As described in Table \@ref(tab:timelygls),  we believe that implementing CHEMs that are transparent, reusable and updatable (TRU) can enable modellers to meet their ethical obligations. We therefore selected these model attributes to use as the basis for deriving ethical assessment criteria.

@@

Transparency has been recommended as a core criterion for assessing ethical public health modelling practice [@10.3389/fpubh.2017.00068]. Guidance on transparency in health economic modelling recommended that model code and data should be clearly documented, potentially with different versions for technical and non-technical users [@Eddy2012]. Notably, the same guidelines, published over ten years ago, did not include recommendations on sharing model code and data. However, more recent guidance recommends publicly dissemination of healthcare model artefacts using online repository services [@Erdemir2020]. Repositories such as Zenodo [@Zenodo2013] and Dataverse [@Dataverse2007] provide persistent storage solutions that generate a Digital Object Identifier (DOI) for each code and data collection. An essential component of quality assuring health economic models is verification - ensuring that calculations are correct and consistent with model specifications [@techver2019]. The extensiveness of verification checks in models implemented as software projects can be reported using the concept of code coverage [@ERICWONG2010188] - the proportion of model code that has been explicitly tested. Tests should ideally combine both unit tests (to verify that small, isolated sections of code produce the correct output when run independently) and acceptance tests (to verify that the correct output is produced when multiple code components are run together to perform tasks that meet core user-requirements [@martin2003agile]). The nature and extent of individual model authorship contributions can become unclear when models are implemented over longer time-frames with a large and changing group of collaborators [@thompson2022escape]. This issue can be addressed by use of online repository services such as GitHub [@github2007], that provide citation tools and can transparently record all individual code contributions to a modelling project over its lifecycle.

Assessment criteria for a transparent CHEM:

-	T1: All model code, non-confidential data and testing procedures and outcomes are available in open access repositories.

-	T2: It is easy to see who developed and tested each part of the model and to identify the modelling team’s assumptions, judgments and theories about model development and use.

Making a CHEM’s code, data and documentation publicly available is helpful but insufficient for promoting model re-use. The choices that CHEM developers make about model implementation and licensing will also shape who can use a model and for what purposes. Using open-source development platforms and licenses can aid both generalizability and transferability. Compared to using commercial modelling software, authoring CHEMs in an open-source language like R [@RCORE2022] makes it easier to store model algorithms and data in distinct files and locations (as opposed to hard coding - embedding data such as parameter values into source code) which facilitates selective modification of model components. This benefit can be further enhanced if model developers adopt a modular approach, in which a model is constructed from multiple reusable and replaceable sub-models (modules) [@pan2021modular]. To grant permissions to others to use, test and adapt models and their components, health economists can avail of two broad categories of open source licensing options. Some guidance strongly recommends the use of permissive licensing [@Wilson_2017] that provides users with great flexibility as to the purposes (including commercial) for which content can be re-used. An alternative approach is to use copyleft licenses [@copyleft2022] that can require content users to distribute any derivative works they create under similar open-source arrangements.

Assessment criteria for a reusable CHEM:

-	R1: Model code and data are implemented to facilitate both generalizability and transferability.

-	R2: Terms of use allow anyone to reuse model code and non-confidential data, in whole or in part, without charge, and for purposes that include the creation of derivative works.

To remain valid for longer, models should be continually updated and refined as new evidence emerges and healthcare systems evolve [@Jenkins2021; @@garcia2021cost]]. Ensuring that a model is regular reviewed to identify and implement required improvements is a receommended defence against model validity decay [@calder2018computational]. Key enablers of sustainable maintenance of open source research software are committed, adequately resourced core development team and active user community [@info:doi/10.2196/20028]. Currently, the core development team for a CHEM will be typically be funded to produce a project end-point deliverable whose specifications are well defined early in the project. For more complex and multi-purpose CHEMs, particularly those designed to be incorporated into decision support systems, it may be better for development teams to adopt Agile Software Development [@beck2001manifesto], an approach that has been recommended for complex public health software projects [@10.3389/fpubh.2022.899874]. An Agile model will be less clearly specified in the initial project plan, but will instead continually develop in response to the requirements and feedback of users, who are provided with an initial, simplified working version of the model at the earliest feasible opportunity. Online communities can be an efficient means of engaging model users in testing each version of a model, identifying issues and suggesting improvements. Services such as GitHub [@github2007] provide collaborative code development tools [@MERGEL2015464] that help elicit, integrate and reconcile contributions from multiple contributors and to ensure each update is uniquely identifiable and retrievable. It is important that verification checks are rerun with each model update, a task that can be automated using the software development practice of Continuous Integration [@CI2017].  The risk of model revisions having unintended consequences for third party users can be mitigated through the use of deprecation conventions [@zhou2016api] that take an informative and staged approach to retiring outdated model code and data.

Assessment criteria for an updatable CHEM:

-	U1: Resources and infrastructure are in place to support sustained development, testing, maintenance and version control of a model in collaboration with model users.

-	U2: Each new release of a model is retested, with changes implemented to minimize disruptions for existing model users.

### 2023-04-19 Markdown/Manuscript/Child_RMDs/MS_Imp_Toolkit.Rmd
draft for selected co-author circulation
first pass and updated bib

We developed a toolkit to help us develop and use MOSHEM modules, datasets and analyses that meet the standards listed in Table \@ref(tab:timelygls). The toolkit is comprised of online repositories and software.

## Repositories <!-- Existing ref -->
We created a GitHub organisation (a collection of code repositories) where all our development code is stored and version controlled [@ready4gh2022]. We configured the repositories in our GitHub organisation to use GitHub actions to support continuous integration. Some of the continuous integration checks we have defined assess each library's compliance with policies specified by the Comprehensive R Archive Network (CRAN) [@CRAN2022], to which we plan submitting future production releases. To track our code coverage, we linked our GitHub organisation to an account we established at codecov [@codecov_2022].

We enabled GitHub Pages in each repository we used for code library development to facilitate the creation and hosting of library documentation websites. We also developed a consolidated and versioned project documentation website [@rfwn2022] using the Hugo framework [@hugo_2023] and Docsy theme [@docsy_2023]. To host that website we established an account with Netlify [@netlify_2023] and linked that service to our GitHub organisation so that the website would automatically update whenever the source code in its GitHub repository was edited.

We also created a Zenodo community [@ready4zen2022] - a collection of permanent, uniquely identified repositories. We then linked our Zenodo community and GitHub organisation so that every time we specify a version of code in one of our GitHub repositories as a "release", a copy of that code is automatically created on Zenodo with a DOI. Finally, to manage model datasets, we created a dedicated collection within the Harvard Dataverse installation [@ready4dv2022].

## Software <!-- Novel ref -->
We created six development version R libraries to help us author model modules, supply those modules with data and implement reproducible modelling analyses. The six libraries, their primary focus, the standards they support and the third-party packages they depend on are summarised in Table \@ref(tab:cpkgs).

A library called ready4 [@ready42022] defines a template module (using R's S4 class system) from which all model module data structures will inherit features and a novel syntax for attaching algorithms to those data structures. The ready4 library also contains tools for retrieving web based information on model modules, datasets and analysis programs and for partially automating updates to the project documentation website.

Three R libraries are designed to standardise and partially automate workflows for authoring new model modules. The ready4pack library [@ready4pack2022] is designed to integrate with our GitHub organisation and provides tools for authoring module libraries that are:

- documented (with a website, a manual itemising selected contents and a manual itemising all contents);

- licensed (using GNU GPL-3 [@GNUGPL2022] by default);

@@

- quality assured (each update triggers continuous integration workflows, including any unit tests created by module library authors).

The ready4pack library depends on two other module authoring libraries. Methods from the ready4fun library [@ready4fun2022] are used to verify that functions for implementing module algorithms are written in a consistent house style. That standardised format is then used by ready4fun methods to automatically generate basic documentation for each function. Methods from the ready4class [@ready4class2022] library are used to streamline and standardise the authoring of module data structures and the linking of methods to these data structures. Like ready4fun, the ready4class library uses standardised code implementation to automatically generate basic documentation for each module data structure.

The ready4use library  [@ready4use2022] contains modules for ingesting model datasets from online repositories (hosted on a Dataverse installation or on GitHub), labelling model datasets and sharing model datasets via online repositories. The ready4show library [@ready4show2022] contains tools to help author analyses programs that are either self-documenting or which trigger the creation of a scientific summary.

When used in conjunction with toolkit repositories, the six R libraries provide support for implementing 17 out of 20 framework standards (Table \@ref(tab:timelygls)). Standards relating to safe dissemination of statistical models (R8), user-interface development (R9) and deprecation conventions (U4) are better met through using existing third party R libraries. Preparing statistical models for dissemination can be accomplished with standard R data management tools like the dplyr [@dplyr2022] and purrr [@dplyr2022] libraries. User-interfaces are typically developed with the shiny [@shiny2022] library, for which a tutorial aimed at health economists is available [@SmithR2020]. The library lifeycle [@lifecycle2021] provides tools for R developers to consistently deprecate their code.

@@

\blandscape

```{r, child = child_docs_ls$three, echo=FALSE}
```

\elandscape

>>>>

To support the collaborative development of CHEMs that meet TRU assessment criteria, we have created a prototype software framework called ready4. We have designed this framework as a lightweight extension to existing software development infrastructure that could be used to implement CHEMs that are both open-source (for transparency and reusability) and modular (for reusability and updatability).

## 	Framework libraries 
To work within the popular open-source programming environment R [@RCORE2022], the ready4 software framework is implemented as six development version R code libraries. The R libraries collectively provide model developers with tools to combine R's functional and object-oriented programming paradigms [@10.1214/13-STS452] in authoring CHEMs, supplying those CHEMs with data and using CHEMs to implement reproducible modelling analyses. The six novel libraries and the preexisting third-party R libraries they depend on are summarised in Table \@ref(tab:cpkgs).

Standardization and interoperability are core requirements of implementing a modular approach. Model modules need to be able to share inputs and outputs with each other and to be run as independent models [@barros2023empowering]. Each CHEM module authored with the ready4 framework will include both a data structure (specifying the required properties of data that can validly be supplied to a module) and a set of algorithms (specifying the operations that can be performed on data contained in a module instance). The foundational framework library, called ready4, defines a template module data structure (using R’s S4 class system) from which all CHEM module data structures will be created and a novel syntax that enable module algorithms to be consistently named. The ready4 library also contains tools for retrieving web based information on model modules, datasets and analysis programs authored with the framework and for partially automating updates to a project documentation website.

Three R libraries are designed to help standardize workflows for authoring, documenting, testing and disseminating new model modules. The ready4pack library is designed to integrate with GitHub and provides tools for authoring model modules and desseminating them as themed bundles in R libraries libraries that are:

-	documented (with a website, a PDF manual itemising selected contents and a PDF manual itemising all contents);

-	licensed (using the copyleft GNU GPL-3 [66] by default);

@@

-	quality assured (each update triggers continuous integration workflows, including any acceptance and unit tests created by module library authors).

The ready4pack library depends on two other module authoring libraries. Writing model algorithms as collections of functions (short, self-contained and reusable software routines that each perform a discrete task), has been recommended as good practice for scientific computing [@Wilson_2017].  The ready4fun library contains tools for authoring functions in a consistent house style that automatically generates basic documentation for each function. Functions to implement model algorithms can be associated with a module via a special type of function called a method. Tools from the ready4class library can help streamline and standardise the authoring of module data structures and their associated methods and to automatically generate basic documentation for each module.

The ready4use library contains tools for supplying model modules with data stored in online repositories (hosted on a Dataverse installation or on GitHub), labelling these datasets and then sharing them via online repositories. The ready4show library contains tools to help author R Markdown programs that combine model modules and datasets to undertake analyses. These programs are either self-documenting (code is easy to understand and integrated with plain English explanations of what it does) or trigger the creation of separate documents (e.g. a scientific manuscript).

@@

##	Framework integration with online services 
Our software framework needs to be used in conjunction with a number of online services. To facilitate its application to our youth mental health project, we established and configured accounts with these required services (see Availability of data and materials).

We created a GitHub organisation (a collection of code repositories) where all code (libraries, programs and sub-routines) that we author with the framework is stored and version controlled. We configured individual repositories in our GitHub organisation that are used for code library projects to use GitHub actions to implement continuous integration. By default, code libraries authored with our framework will use continuous integration to assess compliance with policies specified by the Comprehensive R Archive Network (CRAN) [@CRAN2022]. To track our code coverage, we linked our GitHub organisation to an account we established at codecov [@codecov_2022]. To facilitate the creation and hosting of documentation websites, we enabled GitHub Pages in each repository we used for code library development. 
We also created a Zenodo community - a collection of permanent, uniquely identified repositories. We then linked our Zenodo community and GitHub organisation so that every time we specify a version of code in one of our GitHub repositories as a “release”, a copy of that code is automatically created on Zenodo with a DOI. Finally, to manage model datasets, we created a dedicated collection within the Harvard Dataverse installation.

### 2023-04-19 Markdown/Manuscript/Child_RMDs/MS_Intro_Short.Rmd
first pass and updated bib

<!-- `r ifelse(is.na(params$X@background_1L_chr),"",params$X@background_1L_chr)`  -->
Computational models, particularly those addressing economic topics, have become essential tools for health policy development [@dakin2015influence; @Erdemir2020]. Although influential and widely used, health economic models typically have a number of limitations that at best restrict their usefulness and which in some cases may facilitate harmful misuse.

There is a strong case for making health economists more accountable for the appropriate use and social acceptability of their models. Users of models should be able to assess their adequacy for a particular purpose [@thompson2019escape].  This goal is difficult to achieve in the current context of the poor reproducibility [@Jalali2021; @McManus2019; @Bermejo2017] and frequently insufficient validation [@Ghabri2019] of health economic models. Many health economic models are released with undeclared errors [@Radeva2020]. The value judgments that strongly shape health economic analyses are rarely made explicit, omissions that may lead to socially unacceptable policy recommendations [@duckett2022journey]. A modelling team's value judgments about what questions to address, the most important features of a system to represent and the weighting of different types of evidence may be poorly aligned with those of the people impacted by decisions informed by model analyses [@thompson2022escape]. As health economic models adopt more sophisticated techniques, the need for accountability grows. More complex models may be more prone to propagation errors [@Saltelli2019] and models designed to address multiple questions should be expected to meet more onerous verification and validation obligations [@Eddy2012; @Feenstra2022]. The nature and extent of individual model authorship contributions may be less clear in models implemented over longer time-frames with a large and changing group of collaborators [@thompson2022escape].

The resource and skills intensive undertaking of modelling could be made more tractable if it was easier for health economists to re-use each other's models [@Arnold2010]. However, as many health economic models are owned by pharmaceutical companies and consultancies, commercial considerations can limit the reuse of models and their constituent code and data [@Feenstra2022]. Legal and ethical issues such as privacy and confidentiality requirements further limit public release of some model artefacts [@Pouwels2022]. Transferring a health economic model developed for one jurisdiction for application in another typically involves retaining some features and updating others [@barbieri2010international]. However, models do not always make implementation choices (for example, use of concepts with standardised meanings across jurisdictions<!-- [@garcia2021cost] -->, supplying context specific data as replaceable data-packs and distributing source code under licenses that allow derivative works) that facilitate model transferability.

Health economic models should be updated and refined as new evidence emerges and decision contexts change [@Jenkins2021], but this occurs infrequently [@Sampson2017]. Funding for health economic modelling projects rarely extend to provision of medium term support for model updates and improvements. The career trajectories of health economists can also mitigate against adequate maintenance of a model, it being relatively common for model authors to have moved on from the team that owns the model and / or from working on the health condition for which the model was developed.

A potential strategy for improving model accountability, reusability and updatability is to make models both modular and open-source.

A modular model is constructed from multiple reusable and replaceable sub-models (modules) [@pan2021modular]. Modules can share inputs and outputs with other modules or can be run as independent models [@barros2023empowering]. Advantages of modular models include feasibility (large projects are broken into smaller tasks, with each component independently developed and tested) and flexibility (making it easier to selectively replace or update specific parts of a model and to scale up or down the level of granularity) [@pan2021modular]. Modular approaches are currently being used to facilitate the development of complex computational models in disciplines such as biology [@pan2021modular], neuroscience [@frazier2022robust] and ecology [@barros2023empowering]. In health economics the related and enabling concept of reference models has been recommended [@Afzali2013], but peer reviewed studies describing modular health economic models remain relatively rare, though examples exist in infectious disease [@Trauer2017] and cardiology [@Urach2013].

Modular models provide an opportunity for multiple modelling teams to contribute to, test and reuse models. To enhance this capacity, modular models may be implemented as open source projects that give others liberal permissions to access and use model source code and data [@pan2021modular; @barros2023empowering; @frazier2022robust]. Although there appears to be in principle support from many health economists for greater use of open source health economic models (OSHEMs)  [@Pouwels2022], actual implementations are rare [@Emerson2019; @Michalczyk2018; @Feenstra2022]. Barriers to adoption of OSHEMs include concerns about intellectual property, confidentiality, model misuse and the resources required to support open source implementations [@Pouwels2022; @Wu2019]. Adherence to good practice guidance is an essential requirement for healthcare modelling [@Erdemir2020], but guidelines for implementing OSHEMs remain scarce, piecemeal and need improving [@Sampson2019].

Our interest in modular and open source source approaches developed when we began seeking am appropriate framework for undertaking and validly synthesising diverse types of economic research in mental health. Mental disorders impose high health, social and economic burdens worldwide [@RN8;@GBD2019]. Much of this burden is potentially avertable [@RN25], but poorly financed and organised mental health systems are ill-equipped for this challenge [@RN22;@RN23]. A substantial economic literature already exists to assess the affordability and value for money of mental health interventions [@RN34]. This economic evaluation work is an essential prerequisite for improving allocative efficiency in mental health, but could be of greater value to systems planners if integrated with a broader program of economic research.

We are developing ready4 (https://www.ready4-dev.com), a modular OSHEM (MOSHEM) on the mental health of young people aged 12 to 25. We are initially applying ready4 to four of the twelve domains of health economics identified by Wagstaff and Culyer [@wagstaff2012four]:
<!-- There is therefore an opportunity for health economists to undertake mental health research of significant potential public benefit. -->

- *health and its value* (our projects: utility mapping models);

- *determinants of health and ill-health* (our projects: models for creating rich and representative synthetic household populations from survey data);

- *demand for health and health care* (our projects: spatial epidemiology and help-seeking choice models); and

- *supply of health services* (our projects: a model of primary mental health care services).

Once these projects are completed, our aim is to flexibly combine these models to answer questions in two additional Wagstaff and Culyer domains:

- *efficiency and equity* (our goal: assess the distributional impacts and identify the optimal targeting of care provision); and

- *economic evaluation* (our goal: assess the cost-utility of competing policy options).

To undertake this work we first needed to develop a framework for standardisation. Standardisation is an essential enabler of automation, collaboration, interoperability and transferability in modular and open source approaches. However, we are not aware of any appropriate single source of guidance about what standards should be specified for MOSHEMs. To be deployed in computational models, the standards specified for ready4 will need to be implemented by software, some of which will need to be developed for this specific purpose.

<!-- When planning how to undertake and synthesise this program of work, we became concerned with issues relating to accountability, reusability and updatability. If mistakes are common in individual health economic modelling projects [@Radeva2020], then combining multiple projects may be particularly error prone. The risk of errors may also depend on whether a model is too simple or too complex [@Saltelli2019], so the level of granularity appropriate for one model application may be ill-suited to another. Undertaking a multiannual program of interdependent work raises the risk that models developed in the first phase of a project may need to be updated by the time the project is approaching conclusion. -->

<!-- These concerns led us to seek the MOSHEM approach that provides the transparency to make it easier for errors to be uncovered, the adaptability to fix these errors and the flexibility to facilitate interchangeability between representations of different levels of granularity.  -->

>>>>

<!-- `r ifelse(is.na(params$X@background_1L_chr),"",params$X@background_1L_chr)`  -->

Health economics is a discipline concerned with problems that arise due to scarce resources, such as how to value health and healthcare, allocate healthcare budgets and configure health services [@wagstaff2012four].  In seeking to solve these problems, health economists typically use models which are simplified and selective representations of systems that are believed to influence human health. These representations can be described in words and pictures (a conceptual model), in equations (a mathematical model) or in computer code (a computational model). The predictions reported in health economic studies are typically generated by the execution of a computer program which applies a computational model to compatible data inputs (e.g., parameter values) and performs a sequence of numeric calculations.

Computational models are now widely used to inform health policy and system design [@dakin2015influence; @Erdemir2020; @LO2022100570; @Christene006827]. This level of influence has concomitant ethical responsibilities for model developers that are often poorly understood and inadequately fulfilled [@10.3389/fpubh.2017.00068; @pliakos2021ethics; @thompson2022escape; @calder2018computational].

Computational health economic models (CHEMs) can be implemented using specialized commercial software or authored as bespoke software projects in a programming language. Advantage of commercial modelling tools are simplicity and ease of use, but a software development approach facilitates development of models that may be more realistic, transparent, reusable and adaptable [@incerti2019r]. For CHEMs that are implemented as software projects, a major early decision is selection of an appropriate software framework. A software framework is a shared common technology used by developers to collaboratively author software and which is not typically visible to software end-users [@myllarniemi2018development]. Advantages of using software frameworks include facilitating code reuse and extension, promoting good programming practice and the capability to provide enhanced functionality and performance without additional effort by developers [@edwin2014software]. However, software frameworks can be challenging and time consuming to create [@edwin2014software] and then difficult for others to learn, often requiring developers to undergo specialist training [@myllarniemi2018development]. There is also a risk that a software framework may become excessively complex over time [@edwin2014software].

We are developing a model to explore multiple economic questions relating to the mental health of young people aged 12 to 25.  When making choices about how to implement this model computationally, we wished to facilitate ethical development and use. However, we are not aware of any software framework for implementing CHEMs that adhere to explicitly stated ethical requirements.

### 2023-04-19 Markdown/Manuscript/Child_RMDs/MS_Results.Rmd
first pass and updated bib

# Guidelines for OSHEMs
```{r child="../Child_RMDs/MS_Imp_Guidelines.Rmd", echo=FALSE}
```

# Modelling Toolkit
```{r child="../Child_RMDs/MS_Imp_Toolkit.Rmd", echo=FALSE}
```

>>>>

## Ethical requirements and assessment criteria
```{r child="../Child_RMDs/MS_Imp_Guidelines.Rmd", echo=FALSE}
```

## Software framework
```{r child="../Child_RMDs/MS_Imp_Toolkit.Rmd", echo=FALSE}
```
